---
title: "Code"
output: github_document
---

# Income Data, cell-level

## Kummu (2018) - Gridded global datasets for Gross Domestic Product and Human Development Index over 1990-2015

Source: https://doi.org/10.5061/dryad.dk1j0

Peri (2019) uses cell-level data for income to split observations into 4 income groups (poor, mid-poor, mid-rich, rich). However, two drawbacks:

- low frequency, 10 years intervals between 1980 and 2100
- low resolution, 0.5° (= 30 arc-min, https://www.inchcalculator.com/convert/degree-to-arcminute/)

Kummu (2018) has built a finer global gridded dataset for HDI, GPD, and GDP per capita, at PPP and USD2011, with:

- higher frequency, yearly over 1990-2015
- higher resolution, 0.083° (= 5 arc-min, 6 times more detailed)

Goal: distribute cells based on their GPD per capita into 4 percentiles (25th, 50th, 75th). To do so:

1) Long-run income average for all cells, 2000-2015
2) Assign percentiles

Making income percentile at cell-level and not using country-level data (eg. from World Bank) allows me to keep taking advantage of the fine spatial resolution, capturing within-country variation.

Handling NetCDF files in R:
- https://rdrr.io/cran/ncdf4/man/
- https://pjbartlein.github.io/REarthSysSci/netCDF.html

**RESAMPLING**! Climate data are in 6 arcmin (0.1°). Therefore I need to resample Income data first, from 5 arcmin (0.083°) to 6 arcmin! Loss in resolution is not drastic and the aggregation helps in reducing data complexity. 

```{r}
library(terra)
library(ncdf4)

filepath_GDP_per_capita_PPP_1990_2015_v2 <- "../data/GDP_per_capita_PPP_1990_2015_v2.nc"
r_gdp <- rast(filepath_GDP_per_capita_PPP_1990_2015_v2)

r_gdp
```

```{r}
# Select years 2000–2015
r_gdp_sel <- r_gdp[[11:26]]

# Compute mean across years
r_gdp_mean <- mean(r_gdp_sel, na.rm = TRUE)

# Template at 0.1° (6 arcmin) resolution
r_template <- rast(ext(r_gdp_mean), resolution = 0.1, crs = crs(r_gdp_mean))

# Resample to 0.1°
r_gdp_resampled <- resample(r_gdp_mean, r_template, method = "average")

# Convert to data.frame
gdp_avg <- as.data.frame(r_gdp_resampled, xy = TRUE, na.rm = TRUE)
names(gdp_avg)[3] <- "mean_gdp"
```


### Assign Country Code

```{r}
library(rnaturalearth)
library(sf)
library(FNN)

# Get world countries
world <- ne_countries(scale = "medium", returnclass = "sf")

# Direct spatial join
coords_sf <- st_as_sf(
  gdp_avg[, c("x", "y")], 
  coords = c("x", "y"), 
  crs = 4326
)

country_match <- st_join(coords_sf, world[c("iso_a3_eh")], join = st_intersects)
gdp_avg$country_code <- country_match$iso_a3_eh

# For NA values (border/water cells), assign nearest country
na_mask <- is.na(gdp_avg$country_code)

if(sum(na_mask) > 0) {
  # Get country centroids for nearest neighbor matching
  country_centroids <- st_centroid(world)
  country_coords <- st_coordinates(country_centroids)
  
  # Get coordinates of NA points
  na_coords <- as.matrix(gdp_avg[na_mask, c("x", "y")])
  
  # Find nearest country for each NA point
  nn_idx <- get.knnx(country_coords, na_coords, k = 1)$nn.index[, 1]
  
  # Assign nearest country codes to NA values
  gdp_avg$country_code[na_mask] <- world$iso_a3_eh[nn_idx]
}
```

Cleaning

```{r}
# Remove Greenland
gdp_avg <- gdp_avg[gdp_avg$country_code != "GRL",]

# Remove small islands near Iceland
gdp_avg <- gdp_avg[
  !(gdp_avg$x > -45 & gdp_avg$x < 0 & gdp_avg$y > 67 & gdp_avg$y < 90)
  ,]

# Remove outliers > 120k
gdp_avg <- gdp_avg[gdp_avg$mean_gdp < 120000,]

# Remove without country code
gdp_avg <- gdp_avg[gdp_avg$country_code != "-99",]
```

```{r}
library(ggplot2)

ggplot(gdp_avg, aes(x=x, y=y, fill=mean_gdp)) +
  geom_raster() +
  coord_fixed()
```


Let's see the distribution of the average GDP per capita.

```{r}
# Quartiles
p <- quantile(gdp_avg$mean_gdp, probs = c(0.25, 0.5, 0.75))

# Histogram
ggplot(gdp_avg, aes(mean_gdp)) +
  geom_histogram() +
  geom_vline(xintercept = p[1], color="red") +
  geom_vline(xintercept = p[2], color="red") +
  geom_vline(xintercept = p[3], color="red") +
  annotate("text", x = p[1], y = 10000, label = "25th", vjust = -2, color = "red") +
  annotate("text", x = p[2], y = 10000, label = "50th", vjust = -2, color = "red") +
  annotate("text", x = p[3], y = 10000, label = "75th", vjust = -2, color = "red") +
  scale_x_continuous(labels = scales::label_comma(), breaks = seq(0, 125000, by=25000))
```

Let's split observations in quartiles. In this way, we can later group Net-Migration and Climate data in GDP clusters.

```{r}
# Cut into quartiles
gdp_avg$gdp_quartile <- cut(
  gdp_avg$mean_gdp,
  breaks = c(-Inf, p, Inf),
  labels = c("Q1", "Q2", "Q3", "Q4"),
  include.lowest = TRUE
)
```

Let's plot average GDP per capita.

```{r}
ggplot(gdp_avg, aes(x=x, y=y, fill=mean_gdp)) +
  geom_raster() +
  coord_fixed() +
  scale_fill_viridis_c(name = "GDP per capita", limits = c(0,80000), oob = scales::squish)
```

Let's plot quartiles on the map. It's visible how using cell-level data, instead of country-level, we have within-country variation (and not corresponding to regional administrative borders).

```{r}
ggplot(gdp_avg, aes(x=x, y=y, fill=gdp_quartile)) +
  geom_raster() +
  coord_fixed()
```

Some Summary Stats

```{r}
summary(gdp_avg$mean_gdp)
```

Number of cells: 1,5 mln

- 360° / 0.1° = 3600 cells lon
- 180° / 0.1° = 1800 cells lat
- +- 70% of the world is water

```{r}
length(gdp_avg$mean_gdp)
```

Number of cell in each income group.

```{r}
table(gdp_avg$gdp_quartile)
```

Info ORIGINAL raster

```{r}
ext(r_gdp_mean)
```

```{r}
res(r_gdp_mean)
```

Info RESAMPLED raster

```{r}
ext(r_gdp_resampled)
```

```{r}
res(r_gdp_resampled)
```

This is how the GRID has changed after resampling. 

```{r}
# Area
region <- ext(10, 11, 50, 51) # lon_min, lon_max, lat_min, lat_max

# Crop
r_gdp_original_crop <- crop(r_gdp_mean, region) # original
r_gdp_resampled_crop <- crop(r_gdp_resampled, region) # resampled

# To data frames
df_original <- as.data.frame(r_gdp_original_crop, xy = TRUE, na.rm = TRUE)
df_resampled <- as.data.frame(r_gdp_resampled_crop, xy = TRUE, na.rm = TRUE)

ggplot() +
  # Original
  geom_tile(data = df_original, aes(x = x, y = y), fill = "black", alpha = .5, color = "black", linewidth = 1) +
  
  # Resampled
  geom_tile(data = df_resampled, aes(x = x, y = y), fill = NA, color = "red", linewidth = 1) +
  
  coord_fixed()
```

Let's see the VALUEs.

```{r}
library(ggnewscale) # two color scales

ggplot() +
  # Original
  geom_tile(data = df_original, aes(x = x, y = y, fill = mean)) +
  scale_fill_viridis_c(name = "GDP original", option = "magma", alpha = .8) +
  
  ggnewscale::new_scale_fill() +  # Enables second fill scale
  
  # Resampled
  geom_tile(data = df_resampled, aes(x = x, y = y, fill = mean), alpha = .5, color = "black", linewidth = .5) +
  scale_fill_viridis_c(name = "GDP resampled", option = "viridis", alpha = 0.6) +
  coord_fixed()
```

Let's export our final dataset in csv.

```{r}
write.csv(gdp_avg, "income-gdp-quarter.csv", row.names = FALSE)
```

EXPORT RASTER DATASET FOR INCOME

```{r}
library(terra)

# 1. Convert gdp_quartile to numeric
gdp_avg$gdp_quartile_num <- as.numeric(gdp_avg$gdp_quartile)

# 2. Create two separate rasters from the dataframe
r_mean_gdp <- rast(gdp_avg[, c("x", "y", "mean_gdp")], type = "xyz", crs = crs(r_gdp_resampled))
r_quartile <- rast(gdp_avg[, c("x", "y", "gdp_quartile_num")], type = "xyz", crs = crs(r_gdp_resampled))

# 3. Combine into a multi-layer raster (a "SpatRaster" with 2 layers)
r_income_stack <- c(r_mean_gdp, r_quartile)
names(r_income_stack) <- c("mean_gdp", "gdp_quartile")

# 4. Export to GeoTIFF
writeRaster(r_income_stack, filename = "income-gdp-stack.tif", overwrite = TRUE)
```

