---
title: "Code"
output: github_document
---

# Net-Migration Data, cell level

## Niva (2023) - World’s human migration patterns in 2000–2019 unveiled by high-resolution data

Source: https://doi.org/10.1038/s41562-023-01689-4

Code References:

- (Context) Spatial Data Science: With Applications in R: https://doi.org/10.1201/9780429459016
- (Code) R as GIS for Economists: https://tmieno2.github.io/R-as-GIS-for-Economists-Quarto/

Let's import the dataset produced by Niva (2023):

- Data: https://zenodo.org/records/7997134
- Gridded global Net-Migration (per 1000 people)
- Yearly 2000-2019 (20 years)
- High-resolution, 5 arc-min / 0.083° https://www.inchcalculator.com/convert/degree-to-arcminute/

How the authors calculate the Net-Migration rate: https://github.com/mattikummu/global_net_migration/blob/main/3_net_migration_setup/1_net_migration_setup.R

**RESAMPLING**! Climate data have resolution of 6 arcmin (0.1°). Therefore I need to resample migration data from 5 arcmin (0.083°) to 6 arcmin.

```{r}
library(terra)
library(tidyr)
library(ggplot2)
library(dplyr)
library(data.table)

# Import Net-Migration data, 2000-2019 yearly (20 layers)
filepath_raster_netMgr_2000_2019_annual <- "../data/raster_netMgr_2000_2019_annual.tif"
r_original <- rast(filepath_raster_netMgr_2000_2019_annual)

r_original # 2160 * 4320 * 20 years
```

Original data:

- 2160 × 4320 = ~9.3 million cells per layer
- 7 million NAs per layer = ~75% of cells are NA (oceans/water)
- ~2.3 million valid cells per layer

**RESAMPLING!**

```{r}
# Template at 0.1° (6 arcmin) resolution
r_template <- rast(ext(r_original), resolution = 0.1, crs = crs(r_original))

# Resample to 0.1°
r_resampled <- resample(r_original, r_template, method = "average")

# Convert to data.frame
df_netMgr_2000_2019_annual <- as.data.frame(r_resampled, xy = TRUE, na.rm = TRUE)
```

### Merge with Income-Urban

It's time to merge Income and Net-Migration data.

1. Both datasets use the same coordinate/projection system: EPSG:4326 (WGS84) https://epsg.io/4326. This is reassuring, as allows us to spatially merge the features of the two datasets without having to resample.

2. After RESAMPLING, both datasets have the same resolution: 0.1° / 6 arc-min https://www.inchcalculator.com/convert/degree-to-arcminute/

```{r}
library(FNN)
library(geosphere)  # for accurate geographic distances

income_urban <- fread("urban-income.csv")

# Extract coordinates
coords1 <- as.matrix(df_netMgr_2000_2019_annual[, c("x", "y")])
coords2 <- as.matrix(income_urban[, c("x", "y")])

# Nearest neighbor indices
nn_result <- get.knnx(coords2, coords1, k = 1)
nn_idx <- nn_result$nn.index
nn_dist <- nn_result$nn.dist

# Distances in meters
distances <- distGeo(coords1, coords2[nn_idx, ])

# Combine results
df_merged <- cbind(
  df_netMgr_2000_2019_annual,
  income_urban[nn_idx, ],
  distance_m = distances
)

# Filter to max distance
df_merged <- df_merged[df_merged$distance_m <= 5000, ]
df_merged <- df_merged[, !duplicated(names(df_merged))]
df_merged$distance_m <- NULL
```

### Merge with Temperature

```{r}
temperature <- fread("temperature_panel.csv")

# Transform to wide format with custom column names
temperature_wide <- temperature %>%
  pivot_wider(
    id_cols = c(x, y),
    names_from = year,
    values_from = c(temperature, avg_temp, diff_from_avg),
    names_sep = "_"
  ) %>%
  # Rename columns
  rename_with(~ gsub("temperature_", "temp_", .x)) %>%
  rename_with(~ gsub("avg_temp_", "avg_temp_", .x)) %>%
  rename_with(~ gsub("diff_from_avg_", "diff_from_avg_", .x)) # From 30mln (long) to 1,5mln (wide) rows

# Extract coordinates
coords1 <- as.matrix(df_merged[, c("x", "y")])
coords2 <- as.matrix(temperature_wide[, c("x", "y")])

# Nearest neighbor indices
nn_result <- get.knnx(coords2, coords1, k = 1)
nn_idx <- nn_result$nn.index
nn_dist <- nn_result$nn.dist

# Combine results
df_merged_temp <- cbind(
  df_merged,
  temperature_wide[nn_idx, ],
  distance_m = nn_dist
)

# Filter to max distance
df_merged_temp <- df_merged_temp[df_merged_temp$distance_m <= .5, ]
df_merged_temp <- df_merged_temp[, !duplicated(names(df_merged_temp))]
df_merged_temp$distance_m <- NULL
```

### Merge with Precipitation

```{r}
precipitation <- fread("precipitation_panel.csv")

# Transform to wide format with custom column names
precipitation_wide <- precipitation %>%
  pivot_wider(
    id_cols = c(x, y),
    names_from = year,
    values_from = c(tp, avg_tp),
    names_sep = "_"
  )

# Extract coordinates
coords1 <- as.matrix(df_merged_temp[, c("x", "y")])
coords2 <- as.matrix(precipitation_wide[, c("x", "y")])

# Nearest neighbor indices
nn_result <- get.knnx(coords2, coords1, k = 1)
nn_idx <- nn_result$nn.index
nn_dist <- nn_result$nn.dist

# Distances in meters using geodesic method
#distances <- distGeo(coords1, coords2[nn_idx, ])

# Combine results
df_merged_final <- cbind(
  df_merged_temp,
  precipitation_wide[nn_idx, ],
  distance_m = nn_dist
)

# Optional: Filter to max distance
df_merged_final <- df_merged_final[df_merged_final$distance_m <= .5, ]
df_merged_final <- df_merged_final[, !duplicated(names(df_merged_final))]
df_merged_final$distance_m <- NULL
```

```{r}
rm(precipitation)
rm(precipitation_wide)
rm(temperature)
rm(temperature_wide)
rm(df_netMgr_2000_2019_annual)
rm(r_original)
rm(r_resampled)
rm(r_template)
rm(income_urban)
rm(df_merged)
rm(df_merged_temp)
rm(coords1)
rm(coords2)
rm(nn_dist)
rm(nn_idx)
rm(nn_result)
rm(distances)
gc()
```

From wide to long!

```{r}
library(dplyr)
library(tidyr)

# Reshape the dataset to long format
df_long <- df_merged_final %>%
  pivot_longer(
    cols = `2000`:`2019`,
    names_to = "year",
    values_to = "migration"
  ) %>%
  mutate(
    # temp
    temp = df_merged_final %>%
      select(starts_with("temp_")) %>%
      pivot_longer(cols = everything(), names_to = "year", values_to = "temp") %>%
      mutate(year = gsub("temp_", "", year)) %>%
      pull(temp),
    
    # avg_temp
    avg_temp = df_merged_final %>%
      select(starts_with("avg_temp_")) %>%
      pivot_longer(cols = everything(), names_to = "year", values_to = "avg_temp") %>%
      mutate(year = gsub("avg_temp_", "", year)) %>%
      pull(avg_temp),

    # diff_from_avg_temp
    diff_from_avg_temp = df_merged_final %>%
      select(starts_with("diff_from_avg_")) %>%
      pivot_longer(cols = everything(), names_to = "year", values_to = "diff_from_avg_temp") %>%
      mutate(year = gsub("diff_from_avg_", "", year)) %>%
      pull(diff_from_avg_temp),

    # tp
    tp = df_merged_final %>%
      select(starts_with("tp_")) %>%
      pivot_longer(cols = everything(), names_to = "year", values_to = "tp") %>%
      mutate(year = gsub("tp_", "", year)) %>%
      pull(tp),

    # avg_tp
    avg_tp = df_merged_final %>%
      select(starts_with("avg_tp_")) %>%
      pivot_longer(cols = everything(), names_to = "year", values_to = "avg_tp") %>%
      mutate(year = gsub("avg_tp_", "", year)) %>%
      pull(avg_tp)
  ) %>%
  mutate(year = as.integer(year)) %>%
  select(-matches("^(\\d{4}$|temp_\\d{4}$|avg_temp_\\d{4}$|diff_from_avg_\\d{4}$|tp_\\d{4}$|avg_tp_\\d{4}$)"))

fwrite(df_long, "all_panel_data.csv", row.names = FALSE)

rm(df_merged_final)
gc()
```

NAs/0 values to remove

```{r}
df_long[df_long$migration == 0, ] # 2,7 mln obs
```

```{r}
zeros_by_location <- df_long %>%
  filter(year == 2000 & migration == 0) # 140k obs

ggplot(zeros_by_location, aes(x = x, y = y, fill = "red")) +
  geom_raster() +
  labs(
    title = "Spatial Distribution of Zero Net Migration Areas",
    x = "Longitude",
    y = "Latitude"
  ) +
  coord_fixed()
```

Therefore, let's remove observations with 0.

```{r}
# Remove 0s
df_long <- df_long[df_long$migration != 0,] # From 28mln to 25,2mln obs

summary(df_long$migration)
```

Let's proceed cleaning our dataset, it's time for outliers. I remove outliers at 0.005 and 0.995 percentiles.

```{r}
p <- quantile(df_long$migration, probs = c(0.005, 0.995))

# Filter out outliers (1st, 99th percentile)
df_long <- df_long[
  df_long$migration >= p[1] &
  df_long$migration <= p[2], ] # From 25,2mln to 25mln obs

summary(df_long$migration)
```

Let's plot them together INCOME and MIGRATION

```{r}
# Net-Migration data for 1 year, to be able to plot
oneyear <- df_long[df_long$year == 2000,] # 1,2mln obs

library(patchwork)

p1 <- ggplot(oneyear) +
  geom_raster(aes(x = x, y = y, fill = mean_gdp)) +
  coord_fixed() +
  scale_fill_viridis_c(name = "GDP per capita")


p2 <- ggplot(oneyear) + 
  geom_raster(aes(x = x, y = y, fill = migration)) +
  coord_fixed() +
  scale_fill_viridis_b(name = "Net-Migr/1000pp", limits = c(-50, 50), breaks = seq(-50, 50, 25))

p1 + p2 + plot_layout(ncol = 1)
```

```{r}
# Function to filter
filter_continent <- function(grid, lon_min, lon_max, lat_min, lat_max) {
  subset(grid, x >= lon_min & x <= lon_max & y >= lat_min & y <= lat_max)
}
```

Let's overlap them to check whether they're actually using the same grid after resampling!

First, let's subset our global dataset and select a small area, north-east Italy.

```{r}
# Subset the area of interest
region <- filter_continent(df_long, lon_min = 10, lon_max = 15, lat_min = 44, lat_max = 49)

p3 <- ggplot() +
  geom_tile(data = region, aes(x = x, y = y, fill = migration)) +
  coord_fixed() +
  scale_fill_viridis_c(name = "Avg GDP", option = "magma", alpha = 1)

p4 <- ggplot() +
  geom_tile(data = region, aes(x = x, y = y, fill = migration)) +
  coord_fixed() +
  scale_fill_viridis_c(name = "Net-Migration/1000pp", option = "viridis")

p3 + p4 + plot_layout(ncol = 1)
```

### Merge with Population // outdated

```{r, eval=FALSE}
# Fast Nearest Neighbor
library(FNN)

# Import Population
population <- read.csv("../data/population-2000-2019-01deg.csv")

# Initialize the population column
df_netMgr_2000_2019_annual_long$population <- NA

years <- 2000:2019

# Loop through each year
for(yr in years) {
  message(yr)
  
  # Subset data for current year
  netmgr_yr <- df_netMgr_2000_2019_annual_long[df_netMgr_2000_2019_annual_long$year == yr, ]
  pop_yr <- population[population$year == yr, ]
  
  # Extract coordinates
  netmgr_coords <- as.matrix(netmgr_yr[, c("x", "y")])
  pop_coords <- as.matrix(pop_yr[, c("x", "y")])
  
  # Find nearest neighbors
  nn_idx <- get.knnx(pop_coords, netmgr_coords, k = 1)$nn.index[, 1]
  
  # Assign population values
  df_netMgr_2000_2019_annual_long$population[df_netMgr_2000_2019_annual_long$year == yr] <- 
    pop_yr$population[nn_idx]
}
```


Our Net-Migration dataset is ready. Let's plot the distribution.

Most observations have values very close to 0. For instance, a value of 0.02, considering values are the ratio between net migration and 1,000 people, means that for every 1,000 residents, there was a net gain of only 0.02 people, equivalent to just 2 individuals per 100,000 people. Instead, a value of 1 means that for every 1,000 residents, there was a net gain of 1 person.

These are RATIOs. In a region with a population of 100,000:
- A value of 0.2 would imply a net migration of 20 people (0.2/1000 * 100k)
- A value of 1 would imply a net migration of 100 people (1/1000 * 100k)


```{r}
pp <- quantile(df_long$migration, probs = c(0.25, 0.5, 0.75))

ggplot(data = df_long, aes(migration)) +
  geom_histogram(binwidth = 0.01) +
  xlim(-4, 4) +
    geom_vline(xintercept = pp[1], color="red", alpha=0.5, linetype="dotted") +
  geom_vline(xintercept = pp[2], color="red", alpha=0.5, linetype="dotted") +
  geom_vline(xintercept = pp[3], color="red", alpha=0.5, linetype="dotted") +
  annotate("text", x = pp[1], y = 10000, label = "25th", vjust = -2, color = "red") +
  annotate("text", x = pp[2], y = 10000, label = "50th", vjust = -2, color = "red") +
  annotate("text", x = pp[3], y = 10000, label = "75th", vjust = -2, color = "red")
```


Single observations (cell, year) with POSITIVE Net-Migration = 12 mln

```{r}
df_long[df_long$migration > 0, ]
```

Single observations (cell, year) with NEGATIVE Net-Migration = 12,9 mln

```{r}
df_long[df_long$migration < 0, ]
```


# Table

Fancy table with gt: https://gt.rstudio.com/index.html

TABLE BY INCOME QUARTERS FOR NET-MIGRATION

```{r}
library(gt)

# Summary Statistics
summary_stats <- function(data, variable) {
  data %>%
    summarize(
      Obs = sum(!is.na({{ variable }})),
      Mean = mean({{ variable }}, na.rm = TRUE),
      `St. dev.` = sd({{ variable }}, na.rm = TRUE),
      `25th` = quantile({{ variable }}, 0.25, na.rm = TRUE),
      `75th` = quantile({{ variable }}, 0.75, na.rm = TRUE)
    )
}

# For full sample: make Group character explicitly
summary_migr_full_sample <- summary_stats(df_long, migration) %>%
  mutate(Group = "Full sample")

# For group summaries: convert Group to character after rename
summary_migr_by_group <- df_long %>%
  group_by(gdp_quartile) %>%
  group_modify(~ summary_stats(.x, migration)) %>%
  rename(Group = gdp_quartile) %>%
  mutate(Group = as.character(Group))

# Now bind_rows will work without error
summary_table_migr <- bind_rows(summary_migr_full_sample, summary_migr_by_group) %>%
  select(Group, Obs, Mean, `St. dev.`, `25th`, `75th`)

# Print nice table
gt(summary_table_migr) %>% 
  fmt_number(columns = "Obs", use_seps = TRUE, decimals = 0) %>% 
  fmt_number(columns = c(3,4,5,6), decimals = 2) %>% 
  tab_options(table.width = pct(80)) %>% 
  tab_header(title = "Net migration rates (/1000 people)")
  #gtsave("table_netmigration.png")
```

TABLE BY INCOME QUARTERS FOR TEMPERATURE

```{r}
oneyear <- df_long[df_long$year == 2000, ]

# For full sample: make Group character explicitly
summary_temp_full_sample <- summary_stats(oneyear, avg_temp) %>%
  mutate(Group = "Full sample")

# For group summaries: convert Group to character after rename
summary_temp_by_group <- oneyear %>%
  group_by(gdp_quartile) %>%
  group_modify(~ summary_stats(.x, avg_temp)) %>%
  rename(Group = gdp_quartile) %>%
  mutate(Group = as.character(Group))

# Now bind_rows will work without error
summary_table_temp <- bind_rows(summary_temp_full_sample, summary_temp_by_group) %>%
  select(Group, Obs, Mean, `St. dev.`, `25th`, `75th`)

# Print nice table
gt(summary_table_temp) %>% 
  fmt_number(columns = "Obs", use_seps = TRUE, decimals = 0) %>% 
  fmt_number(columns = c(3,4,5,6), decimals = 2) %>% 
  tab_options(table.width = pct(80)) %>% 
  tab_header(title = "Temperature (°C)")
  #gtsave("table_temperature.png")
```

TABLE BY INCOME QUARTERS FOR PRECIPITATIONS

```{r}
# For full sample: make Group character explicitly
summary_prec_full_sample <- summary_stats(oneyear, avg_tp) %>%
  mutate(Group = "Full sample")

# For group summaries: convert Group to character after rename
summary_prec_by_group <- oneyear %>%
  group_by(gdp_quartile) %>%
  group_modify(~ summary_stats(.x, avg_tp)) %>%
  rename(Group = gdp_quartile) %>%
  mutate(Group = as.character(Group))  # <-- important

# Now bind_rows will work without error
summary_table_prec <- bind_rows(summary_prec_full_sample, summary_prec_by_group) %>%
  select(Group, Obs, Mean, `St. dev.`, `25th`, `75th`)

# Print nice table
gt(summary_table_prec) %>% 
  fmt_number(columns = "Obs", use_seps = TRUE, decimals = 0) %>% 
  fmt_number(columns = c(3,4,5,6), decimals = 3) %>% 
  tab_options(table.width = pct(80)) %>% 
  tab_header(title = "Total Precipitation (m)")
  #gtsave("table_precipitations.png")
```


Replication Peri (2019), Table 2: Net Migration in Rural Locations and Changes in Temperatures


```{r}
library(ggplot2)
library(scattermore)
library(dplyr)

# Subset rural data
rural_cells <- df_long %>% filter(urban == 1)

# Compute slope manually per gdp_quartile
coefs <- rural_cells %>%
  group_by(gdp_quartile) %>%
  summarise(
    slope = coef(lm(migration ~ diff_from_avg_temp))[2],
    intercept = coef(lm(migration ~ diff_from_avg_temp))[1],
    .groups = "drop"
  )

# Create regression lines for plotting
reg_lines <- rural_cells %>%
  group_by(gdp_quartile) %>%
  summarise(
    min_x = min(diff_from_avg_temp, na.rm = TRUE),
    max_x = max(diff_from_avg_temp, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  left_join(coefs, by = "gdp_quartile") %>%
  rowwise() %>%
  do({
    data.frame(
      gdp_quartile = .$gdp_quartile,
      x = seq(.$min_x, .$max_x, length.out = 100),
      y = .$intercept + .$slope * seq(.$min_x, .$max_x, length.out = 100)
    )
  }) %>%
  ungroup()

# Create labels for slope annotations
labels <- coefs %>%
  mutate(label = paste0("Slope: ", round(slope, 3)))

# Plot
ggplot(rural_cells, aes(x = diff_from_avg_temp, y = migration)) +
  geom_scattermore(pointsize = 1, alpha = 0.1) +
  geom_line(data = reg_lines, aes(x = x, y = y), color = "red", inherit.aes = FALSE) +
  geom_text(
    data = labels,
    aes(x = Inf, y = Inf, label = label),
    hjust = 1.1, vjust = 1.5,
    inherit.aes = FALSE
  ) +
  facet_grid(. ~ gdp_quartile) +
  theme_minimal() +
  labs(
    title = "Migration vs. Temperature Deviation in Rural Areas",
    x = "Temperature Deviation from Average (°C)",
    y = "Net Migration per 1,000 People"
  )
```


```{r}
library(dplyr)
library(ggplot2)
library(scattermore)

# Subset urban data
urban_cells <- df_long %>% filter(urban == 4)

# Compute slope manually per gdp_quartile
coefs <- urban_cells %>%
  group_by(gdp_quartile) %>%
  summarise(
    slope = coef(lm(migration ~ diff_from_avg_temp))[2],
    intercept = coef(lm(migration ~ diff_from_avg_temp))[1],
    .groups = "drop"
  )

# Create regression lines for plotting
reg_lines <- urban_cells %>%
  group_by(gdp_quartile) %>%
  summarise(
    min_x = min(diff_from_avg_temp, na.rm = TRUE),
    max_x = max(diff_from_avg_temp, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  left_join(coefs, by = "gdp_quartile") %>%
  rowwise() %>%
  do({
    data.frame(
      gdp_quartile = .$gdp_quartile,
      x = seq(.$min_x, .$max_x, length.out = 100),
      y = .$intercept + .$slope * seq(.$min_x, .$max_x, length.out = 100)
    )
  }) %>%
  ungroup()

# Create labels for slope annotations
labels <- coefs %>%
  mutate(label = paste0("Slope: ", round(slope, 3)))

# Plot
ggplot(urban_cells, aes(x = diff_from_avg_temp, y = migration)) +
  geom_scattermore(pointsize = 1, alpha = 0.1) +
  geom_line(data = reg_lines, aes(x = x, y = y), color = "red", inherit.aes = FALSE) +
  geom_text(
    data = labels,
    aes(x = Inf, y = Inf, label = label),
    hjust = 1.1, vjust = 1.5,
    inherit.aes = FALSE
  ) +
  facet_grid(. ~ gdp_quartile) +
  theme_minimal() +
  labs(
    title = "Migration vs. Temperature Deviation in Urban Areas",
    x = "Temperature Deviation from Average (°C)",
    y = "Net Migration per 1,000 People"
  )
```


