---
title: "Code"
output: github_document
---

# Precipitation data, cell level

## ERA5-Land monthly averaged data

Source: https://doi.org/10.24381/cds.68d2bb30

For precipitation data, I use the ERA5-Land dataset from Copernicus.

Definition of "Total Precipitation": Accumulated liquid and frozen water, including rain and snow, that falls to the Earth's surface. This variable is accumulated from the beginning of the forecast time to the end of the forecast step (= monthly).

Unit measure: meters.

In addition to the others preparatory steps, I need to compute the "cumulated total precipitation" of each year. Making the average over a year of the monthly precipitations is not useful to represent the area.

```{r, eval=TRUE}
library(ncdf4)
library(terra)
library(ggplot2)

# 1) Load data

filename <- "../data/total_precipitation.nc"
nc <- nc_open(filename)

# Load latitude and longitude
lat <- ncvar_get(nc, "latitude") # 1801
lon <- ncvar_get(nc, "longitude") # 3600

lon <- ifelse(lon > 180, lon - 360, lon) # Center Europe
lon_order <- order(lon) 
lon <- lon[lon_order] # Reorder starting from the "new" 0

time <- ncvar_get(nc, "valid_time") # Read time: 12 months * 20 years = 240 months
dates <- as.POSIXct(time, origin = "1970-01-01", tz = "UTC") # Convert from seconds to date
years <- format(dates, "%Y")
years_unique <- unique(years) # 2000-2019

# Read precipitations
tp <- ncvar_get(nc, "tp") # Three dimensions: [lon, lat, time]
tp <- tp[lon_order,,]

nc_close(nc)
rm(nc) # free memory
gc() # free memory

# 2) Create rasters and compute year average

dir.create("tp_year_rasters", showWarnings = FALSE) # Create folder

# Loop years
for (j in seq_along(years_unique)) {
  yr <- years_unique[j]
  print(yr)
  idx <- which(years == yr) # Subset months with year

  month_rasters <- vector("list", length(idx))

  for (k in seq_along(idx)) {
    print(k)

    # Create a raster for each month
    r <- rast(t(tp[,,idx[k]]))
    ext(r) <- ext(min(lon), max(lon), min(lat), max(lat)) # Set Extent
    crs(r) <- "EPSG:4326" # Set CRS WGS84

    # Remove Antarctica: below -60°
    r <- crop(r, ext(min(lon), max(lon), -60, max(lat)))

    month_rasters[[k]] <- r
  }

  # Compute YEAR CUMULATIVE SUM
  yearly_raster <- sum(rast(month_rasters), na.rm = TRUE)
  
  # Name the raster layer with the year
  names(yearly_raster) <- paste0("tp_", yr)
  
  writeRaster(yearly_raster, sprintf("tp_year_rasters/year_%s.tif", yr), overwrite = TRUE)

  rm(month_rasters) # free memory
  gc() # free memory
}

rm(tp)
rm(r)
rm(yearly_raster)
gc()

# 3) Yearly rasters

yearly_files <- list.files("tp_year_rasters", pattern = "\\.tif$", full.names = TRUE)

# Read all rasters into a stack
yearly_rasters <- rast(yearly_files)

r_2000 <- yearly_rasters[[1]]  # First raster (2000)
r_2019 <- yearly_rasters[[20]]  # Last raster (2019)

# Compute the overall average
avg_raster <- mean(yearly_rasters)

# Convert each raster to a dataframe
df_2000 <- as.data.frame(r_2000, xy = TRUE)
df_2019 <- as.data.frame(r_2019, xy = TRUE)
df_avg <- as.data.frame(avg_raster, xy = TRUE)

# Merge dataframes
df <- merge(df_2000, df_2019, by = c("x", "y"), suffixes = c("_2000", "_2019"))
df <- merge(df, df_avg, by = c("x", "y"))
colnames(df) <- c("x", "y", "tp_2000", "tp_2019", "avg_tp")
df$prec_diff <- df$tp_2019 - df$avg_tp

write.csv(df, "total_precipitation_summary.csv", row.names = FALSE)
writeRaster(yearly_rasters, "precipitation_resampled_average.tif")
```


```{r}
library(terra)
library(dplyr)
library(tidyr)

# Read raster
r <- rast("precipitation_resampled_average.tif")
years <- 2000:2019
names(r) <- as.character(years)

# Extract all values and coordinates
vals <- values(r)               # matrix: n_cells x 20 years
coords <- crds(r, na.rm=FALSE)  # matrix: n_cells x 2

# Create a data frame
df <- as.data.frame(vals)
colnames(df) <- as.character(years)
df$x <- coords[, 1]
df$y <- coords[, 2]

# Remove rows that are NA across all years
df <- df[rowSums(is.na(df[as.character(years)])) < length(years), ]

# Compute average for each cell
df$avg_tp <- rowMeans(df[as.character(years)], na.rm = TRUE)

# Pivot to long format with one row per year per cell
df_long <- df %>%
  pivot_longer(
    cols = all_of(as.character(years)),
    names_to = "year",
    values_to = "tp"
  ) %>%
  mutate(
    year = as.integer(year)
  ) %>%
  select(x, y, year, tp, avg_tp) %>%
  arrange(x, y, year)

# Export to CSV
write.csv(df_long, "precipitation_panel.csv", row.names = FALSE)
```

Summary Stats for Long Run Total Precipitation

Max = 1,1m/year // Mean = 26cm/year

```{r}
# Summary Stats

q1_tp <- quantile(df$avg_tp, 0.25)
q3_tp <- quantile(df$avg_tp, 0.75)
avg_tp <- mean(df$avg_tp)

summary(df$avg_tp)
```

Distribution for Long Run Total Precipitation average

```{r}
ggplot() +
  geom_histogram(data = df, mapping = aes(avg_tp), bins=100) +
  scale_x_continuous(limits = c(0, .2))  +
  
  geom_vline(xintercept = q1_tp, color = "red", linetype=2) + 
  annotate("text", x = q1_tp-.01, y=50000, label = "25th", color="red") +
  
  geom_vline(xintercept = avg_tp, color = "red", linetype=2) + 
  annotate("text", x = avg_tp, y=50000, label = "avg", color="red") +
  
  geom_vline(xintercept = q3_tp, color = "red", linetype=2) + 
  annotate("text", x = q3_tp+.01, y=50000, label = "75th", color="red") +
  
  ggtitle("Histogram for Long Run World Temperature Average (2000-2019)")
```

When looking at total rainfall over the year, very wet areas like the Amazon rainforest and dry areas like North Africa are easy to see.

```{r}
library(ggplot2)

ggplot() + 
  geom_raster(data = df, mapping = aes(x = x, y = y, fill = avg_tp)) +
  coord_fixed() +
  scale_fill_viridis_b(name = "m", limits = c(0, .1), breaks = seq(0, .1, .025)) +
  ggtitle("All-time Avg Total Precipitations 2000-2019")
```

```{r}
ggplot() +
  geom_histogram(data = df, aes(x = tp_2000), 
                 fill = "steelblue", alpha = 0.5, bins = 100) +
  geom_histogram(data = df, aes(x = tp_2019), 
                 fill = "tomato", alpha = 0.5, bins = 100) +
  labs(x = "Total Precipitations (m)", y = "Count", title = "Histogram of Total Precipitations: 2000 vs 2019") +
  scale_x_continuous(limits = c(0, 0.2)) +
  guides(fill = guide_legend(title = "Year"))
```


```{r}
summary(df$prec_diff)
```

Let's plot where it rained more or less than then Long Run average.

```{r}
ggplot() + 
  geom_raster(data = df, mapping = aes(x = x, y = y, fill = prec_diff)) +
  scale_fill_viridis_b(limits = c(-.15, .15), breaks=seq(-.15, .15, .15))  +
  coord_fixed() +
  ggtitle("Total Precipitation difference 2019 vs Avg")
```

