---
title: "Code"
output: github_document
---

# Temperature data, cell level

## ERA5-Land monthly averaged data

Source: https://doi.org/10.24381/cds.68d2bb30

For temperature data, I use the ERA5-Land dataset from Copernicus. It provides "2m temperature" at grid cell level, with a resolution of 0.1°, and a monthly frequency from 2000 to 2019.

Net-Migration and GDP data have a resolution of 0.083°. I downscale them to a coarser resolution of 0.1° to match them with climate data.

```{r, eval=FALSE}
library(ncdf4)
library(terra)

# 1) Load data

filename <- "../data/2m_temperature.nc"
nc <- nc_open(filename)

# Load latitude and longitude
lat <- ncvar_get(nc, "latitude") # 1801
lon <- ncvar_get(nc, "longitude") # 3600

lon <- ifelse(lon > 180, lon - 360, lon) # Center Europe
lon_order <- order(lon) 
lon <- lon[lon_order] # Reorder starting from the "new" 0

time <- ncvar_get(nc, "valid_time") # Read time: 12 months * 20 years = 240 months
dates <- as.POSIXct(time, origin = "1970-01-01", tz = "UTC") # Convert from seconds to date
years <- format(dates, "%Y")
years_unique <- unique(years) # 2000-2019

# Read temperature
temp_k <- ncvar_get(nc, "t2m")  # Three dimensions: [lon, lat, time]
temp_c <- temp_k[lon_order, , ] - 273.15  # Reoder and convert from Kelvin to Celsius

nc_close(nc)
rm(nc) # free memory
rm(temp_k) # free memory
gc() # free memory

# 2) Create rasters and compute year average

dir.create("temp_year_rasters", showWarnings = FALSE) # Create folder

# Loop years
for (j in seq_along(years_unique)) {
  yr <- years_unique[j]
  print(yr)
  idx <- which(years == yr) # Subset months with year

  month_rasters <- vector("list", length(idx))

  for (k in seq_along(idx)) {
    print(k)

    # Create a raster for each month
    r <- rast(t(temp_c[,,idx[k]]))
    ext(r) <- ext(min(lon), max(lon), min(lat), max(lat)) # Set Extent
    crs(r) <- "EPSG:4326" # Set CRS WGS84

    # Remove Antarctica: below -60°
    r <- crop(r, ext(min(lon), max(lon), -60, max(lat)))

    month_rasters[[k]] <- r
  }

  # Compute year mean
  yearly_raster <- mean(rast(month_rasters), na.rm = TRUE)
  
  # Global mean temperature for the year
  mean_value <- global(yearly_raster, fun = "mean", na.rm = TRUE)[1,1]
  cat("Year:", yr, "- Mean temp (°C):", round(mean_value, 2), "\n")
  
  # Name the raster layer with the year
  names(yearly_raster) <- paste0("temp_", yr)
  
  writeRaster(yearly_raster, sprintf("temp_year_rasters/year_%s.tif", yr), overwrite = TRUE)

  rm(month_rasters) # free memory
  gc() # free memory
}

rm(temp_c)
rm(r)
rm(yearly_raster)
gc()

# 3) Yearly rasters

yearly_files <- list.files("temp_year_rasters", pattern = "\\.tif$", full.names = TRUE)

# Read all rasters into a stack
yearly_rasters <- rast(yearly_files)

r_2000 <- yearly_rasters[[1]]  # First raster (2000)
r_2019 <- yearly_rasters[[20]]  # Last raster (2019)

# Compute the overall time average
time_avg_raster <- mean(yearly_rasters)

# Convert each raster to a dataframe
df_2000 <- as.data.frame(r_2000, xy = TRUE)
df_2019 <- as.data.frame(r_2019, xy = TRUE)
df_avg <- as.data.frame(time_avg_raster, xy = TRUE)

# Merge dataframes
df <- merge(df_2000, df_2019, by = c("x", "y"), suffixes = c("_2000", "_2019"))
df <- merge(df, df_avg, by = c("x", "y"))
colnames(df) <- c("x", "y", "temp_2000", "temp_2019", "avg_temp")
df$temp_diff <- df$temp_2019 - df$avg_temp # Temp. diff. between last year and long run avg

write.csv(df, "temperature_summary.csv", row.names = FALSE)
writeRaster(yearly_rasters, "temperature_resampled_average.tif")
```

```{r}
library(terra)
library(dplyr)
library(tidyr)

# Read raster (20 layers: one per year)
r <- rast("temperature_resampled_average.tif")
years <- 2000:2019
names(r) <- as.character(years)

# Extract all values and coordinates
vals <- values(r)               # matrix: n_cells x 20 years
coords <- crds(r, na.rm=FALSE)  # matrix: n_cells x 2

# Create a data frame
df <- as.data.frame(vals)
colnames(df) <- as.character(years)
df$x <- coords[, 1]
df$y <- coords[, 2]

# Remove rows that are NA across all years
df <- df[rowSums(is.na(df[as.character(years)])) < length(years), ]

# Compute average temperature for each cell
df$avg_temp <- rowMeans(df[as.character(years)], na.rm = TRUE)

# Pivot to long format with one row per year per cell
df_long <- df %>%
  pivot_longer(
    cols = all_of(as.character(years)),
    names_to = "year",
    values_to = "temperature"
  ) %>%
  mutate(
    year = as.integer(year),
    diff_from_avg = temperature - avg_temp
  ) %>%
  select(x, y, year, temperature, avg_temp, diff_from_avg) %>%
  arrange(x, y, year)

# Export to CSV
#write.csv(df_long, "temperature_panel.csv", row.names = FALSE)
```


```{r}
library(ggplot2)

ggplot() + 
  geom_raster(data = df, mapping = aes(x = x, y = y, fill = avg_temp)) +
  coord_fixed() +
  scale_fill_viridis_c(name = "°C") +
  ggtitle("All-time Avg Temp 2000-2019")
```

Summary Stats for Long Run Temperature average

```{r}
# Summary Stats

q1_temp <- quantile(df$avg_temp, 0.25)
q3_temp <- quantile(df$avg_temp, 0.75)
avg_temp <- mean(df$avg_temp)

summary(df$avg_temp)
```

Distribution for Long Run Temperature average

```{r}
ggplot() +
  geom_histogram(data = df, mapping = aes(avg_temp), bins=100) +
  scale_x_continuous(limits = c(-30, 35), breaks = seq(-30, 35, 5)) +
  
  geom_vline(xintercept = q1_temp, color = "red", linetype=2) + 
  annotate("text", x = q1_temp-2, y=50000, label = "25th", color="red") +
  
  geom_vline(xintercept = avg_temp, color = "red", linetype=2) + 
  annotate("text", x = avg_temp-2, y=50000, label = "avg", color="red") +
  
  geom_vline(xintercept = q3_temp, color = "red", linetype=2) + 
  annotate("text", x = q3_temp-2, y=50000, label = "75th", color="red") +
  
  ggtitle("Histogram for Long Run World Temperature Average (2000-2019)")
```


```{r}
df$temp_diff <- df$`2019` - df$avg_temp

ggplot() + 
  geom_raster(data = df, mapping = aes(x = x, y = y, fill = temp_diff)) +
  coord_fixed() +
  scale_fill_viridis_b(name = "°C", breaks = seq(-2, 2, 1)) +
  ggtitle("Temperature Difference between 2019 and Long-Run Avg")
```

Distribution of the temperature difference between 2019 and the long-run average. The mean difference is +0.5°C, indicating a general warming. The interquartile range (25th–75th percentile) is entirely above zero, suggesting that the majority of observations experienced above-average temperatures in 2019.

```{r}
# Summary Stats

q1_temp <- quantile(df$temp_diff, 0.25)
q3_temp <- quantile(df$temp_diff, 0.75)
avg_temp <- mean(df$temp_diff)

summary(df$temp_diff)
```

```{r}
ggplot() +
  geom_histogram(data = df, mapping = aes(temp_diff), bins=100) +
  scale_x_continuous(breaks = seq(-3, 3, 1)) +
  
  geom_vline(xintercept = q1_temp, color = "red", linetype=2) + 
  annotate("text", x = q1_temp-.2, y=60000, label = "25th", color="red") +
  
  geom_vline(xintercept = avg_temp, color = "red", linetype=2) + 
  annotate("text", x = avg_temp, y=60000, label = "avg", color="red") +
  
  geom_vline(xintercept = q3_temp, color = "red", linetype=2) + 
  annotate("text", x = q3_temp+.2, y=60000, label = "75th", color="red") +
  
  ggtitle("Histogram for Temperature Difference 2019-Avg")
```

Distribution of temperature in 2000 vs. 2019. The 2019 distribution is clearly shifted to the right, indicating generally higher temperatures compared to 2000. However, since these are single-year snapshots, they may reflect anomalous conditions and should be interpreted with caution.

```{r}
ggplot() +
  geom_histogram(data = df, aes(x = `2000`), 
                 fill = "steelblue", alpha = 0.5, bins = 100) +
  geom_histogram(data = df, aes(x = `2019`), 
                 fill = "tomato", alpha = 0.5, bins = 100) +
  labs(x = "Temperature (°C)", y = "Count", title = "Histogram of Temperatures: 2000 vs 2019") +
  scale_x_continuous(limits = c(-35, 35), breaks = seq(-35, 35, 5)) +
  guides(fill = guide_legend(title = "Year"))
```

