---
title: "Code"
output: github_document
---

# Urban-Rural data, cell-level

## Global Human Settlement - Degree of Urbanisation (GHS-SMOD)

Source: https://human-settlement.emergency.copernicus.eu/ghs_smod2023.php

Another fundamental piece of information for my analysis is the level of urbanization of each cell. In addition to the fact that most migrations occur from low-GDP to high-GDP countries, most migrations also originate in rural areas and point to urban areas.

Peri (2019) clusters cells based on the population density distribution for each country, creating four groups: rural, semi-rural, semi-urban, and urban. However, using only population density to categorize cell types is not ideal. For example, population density alone doesn't account for other key factors, such as economic activity or the built-up area in the cell, which are essential for correctly classifying a cell.

To this end, the European JRC have built a dataset to comprehensively cluster cells into 8 categories, following the Degree of Urbanisation (European Commission & Statistical Office). To do so, they use data from two dataset: GHS-POP and GHS-BUILT-S.

- 30: URBAN CENTRE
- 23: DENSE URBAN
- 22: SEMI-DENSE URBAN
- 21: SUBURBAN
- 13: RURAL
- 12: LOW DENSITY RURAL
- 11: VERY LOW DENSITY RURAL
- 10: WATER

To stick to the usual four clusters/quartiles methodology, and to reproduce Peri (2019), I re-classify the categories as follow:

- URBAN: 30, 23
- SEMI-URBAN: 22, 21
- SEMI-RURAL: 13, 12
- RURAL: 11
- Water cells are removed

Additionally, the JRC provides data at five-year intervals from 2000 to 2020. Rather than arbitrarily selecting a single observation (e.g., the oldest), I decide to assign the mode of the data to each cell for more consistent categorization.

```{r}
library(terra)
library(ggplot2)
library(viridis)

# 1. Load the 5 SMOD rasters into a SpatRaster stack
files <- c(
  "GHS_SMOD_E2000_GLOBE_R2023A_4326_30ss_V2_0.tif",
  "GHS_SMOD_E2005_GLOBE_R2023A_4326_30ss_V2_0.tif",
  "GHS_SMOD_E2010_GLOBE_R2023A_4326_30ss_V2_0.tif",
  "GHS_SMOD_E2015_GLOBE_R2023A_4326_30ss_V2_0.tif",
  "GHS_SMOD_E2020_GLOBE_R2023A_4326_30ss_V2_0.tif"
)
stack <- rast(file.path("../data/GHS_SMOD/", files))

# 2. Compute pixel-wise mode across the 5 years
mode <- modal(stack, na.rm = TRUE)

# 3. Reclassify values into your four categories
# 30 → 4 (URBAN)
# 23, 22 → 3 (SEMI-URBAN)
# 21, 13 → 2 (SEMI-RURAL)
# 12, 11 → 1 (RURAL)
# 10 (WATER) → NA

rcl <- matrix(c(
  30, 4,
  23, 4,
  22, 3,
  21, 3,
  13, 2,
  12, 2,
  11, 1,
  10, NA
), ncol = 2, byrow = TRUE)

class <- classify(mode, rcl)

# Define target raster with 0.1 degree resolution
target_res <- 0.1
target_raster <- rast(ext(class), resolution = target_res, crs = crs(class))

# Nearest neighbor keeps the original class values intact by assigning each output cell the value of the closest input cell, preserving the discrete categories
downscaled <- resample(class, target_raster, method = "near")

# Define extent excluding Antarctica
e_no_antarctica <- ext(xmin(downscaled), xmax(downscaled), -60, ymax(downscaled))

# Crop the downscaled raster
downscaled <- crop(downscaled, e_no_antarctica)

df <- as.data.frame(downscaled, xy = TRUE, na.rm = TRUE)
names(df)[3] <- "urban"

# Remove rows where urban == 0
df <- df[df$urban != 0, ]

ggplot() + 
  geom_raster(data  = df, mapping = aes(x = x, y = y, fill = factor(urban))) +
  coord_fixed()
```

MERGE WITH INCOME DATA: CROSS-CROSS

```{r}
library(terra)

# Load the income raster stack
income_stack <- rast("income-gdp-stack.tif")

# Convert income raster to points (this defines our target locations)
income_points <- as.points(income_stack, na.rm = TRUE)

# Extract urban values at income point locations
urban_values <- extract(downscaled, income_points, method = "simple")

# Combine the data
# income_points already contains the coordinates and income data
# urban_values contains the ID and urban classification
df_merged <- as.data.frame(income_points, geom = "XY")
df_merged$urban <- urban_values[,2]

df_merged <- df_merged[!is.na(df_merged$urban),]
df_merged <- df_merged[df_merged$urban > 0,]

# Reorder cols
df_merged <- df_merged[,c(3,4,1,2,5)]
```

ASSIGN COUNTRY CODE

```{r}
library(rnaturalearth)
library(sf)
library(FNN)

# 1. Load world country polygons
world <- ne_countries(scale = "medium", returnclass = "sf")

# 2. Convert df_merged coordinates to sf object
coords_sf <- st_as_sf(
  df_merged[, c("x", "y")], 
  coords = c("x", "y"), 
  crs = 4326
)

# 3. Spatial join: assign ISO country code
country_match <- st_join(coords_sf, world[c("iso_a3_eh")], join = st_intersects)

# 4. Add country code to df_merged
df_merged$country_code <- country_match$iso_a3_eh

# 5. Handle missing values (cells not intersecting any country polygon)
na_mask <- is.na(df_merged$country_code)

if(sum(na_mask) > 0) {
  # Get centroids of countries
  country_centroids <- st_centroid(world)
  country_coords <- st_coordinates(country_centroids)
  
  # Get coordinates of unmatched cells
  na_coords <- as.matrix(df_merged[na_mask, c("x", "y")])
  
  # Find nearest country centroid for each NA point
  nn_idx <- get.knnx(country_coords, na_coords, k = 1)$nn.index[, 1]
  
  # Assign closest country code
  df_merged$country_code[na_mask] <- world$iso_a3_eh[nn_idx]
}
```

SUMMMARY STATISTICS

```{r}
library(ggplot2)

# Define labels
urban_labels <- c("Rural", "Semi-Rural", "Semi-Urban", "Urban")

# Create summary table
urban_summary <- as.data.frame(table(df_merged$urban))
names(urban_summary) <- c("urban", "count")
urban_summary$urban <- factor(urban_summary$urban,
                              levels = 1:4,
                              labels = urban_labels)
urban_summary$share <- 100 * urban_summary$count / sum(urban_summary$count)

# Plot
ggplot(urban_summary, aes(x = urban, y = share, fill = urban)) +
  geom_col() +
  labs(
    title = "Share of Urban Types",
    x = NULL,
    y = "Share (%)"
  ) +
  theme_minimal()
```

```{r}
library(dplyr)
library(ggplot2)

# Define labels for urban classes
urban_labels <- c("Rural", "Semi-Rural", "Semi-Urban", "Urban")

# Calculate mean GDP per urban class
gdp_by_urban <- df_merged %>%
  group_by(urban) %>%
  summarise(mean_gdp = mean(mean_gdp, na.rm = TRUE)) %>%
  mutate(urban = factor(urban, levels = 1:4, labels = urban_labels))

# Plot
ggplot(gdp_by_urban, aes(x = urban, y = mean_gdp, fill = urban)) +
  geom_col() +
  labs(
    title = "Mean GDP by Urban Class",
    x = NULL,
    y = "Mean GDP"
  ) +
  theme_minimal()
```


```{r}
write.csv(df, "urban-income.csv", row.names = FALSE)
```

